---
title: "Getting Started with BiocKB"
author: "Jacques Serizay"
date: "`r Sys.Date()`"
output:
    BiocStyle::html_document:
        toc: true
vignette: >
    %\VignetteIndexEntry{Getting Started with BiocKB}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Introduction

**BiocKB** builds and manages a RAG (Retrieval-Augmented Generation) knowledge
base from Bioconductor documentation. It provides a DuckDB-backed vector store
with chunked, embedded content from package vignettes, man pages, and other
Bioconductor resources.

# Installation

```{r install}
BiocManager::install("BiocKB")
```

# Store lifecycle

The typical workflow follows: fetch → chunk → insert → index → retrieve.

```{r create}
library(BiocKB)
```

## Fetch Bioconductor resources

```{r fetch-pkgs}
base_dir = "~/appascratch/biockb_resources"

## Fetch resources from Bioconductor packages
pkgs_files <- fetch_pkgs_resources(file.path(base_dir, "pkgs"))
pkgs_files <- gsub("pkgs/all_files", "all_files/pkgs", pkgs_files)

## Fetch resources from Bioconductor website
pages <- fetch_bioc_website(base_dir) |> unlist()

## Fetch resources from Bioconductor VIEWS
views <- fetch_bioc_views(base_dir)

## Fetch resources from Package Developer book website
contribs <- fetch_bioc_contributions(base_dir) |> unlist()

## Fetch resources from Bioconductor support site posts
# posts <- fetch_support_posts(base_dir)

all_files <- c(pages, views, contribs, pkgs_files)
saveRDS(all_files, "all_files.rds")
```

## Chunk and insert

```{r insert}
all_files <- readRDS("all_files.rds")
all_files <- c(
    grep("bioconductor/", all_files, value = TRUE) |> grep("_release.md", x = _, invert = TRUE, value = TRUE) |> sample(100),
    grep("biocviews/", all_files, value = TRUE) |> sample(100),
    grep("pkgrevdocs/", all_files, value = TRUE),
    grep("DEXSeq", all_files, value = TRUE)
)

## Chunk all files in parallel
library(future)
plan(future::multicore, workers = 32)
chunks <- furrr::future_map(
    all_files, 
    chunk_bioc_file, 
    base_path = file.path(base_dir, "all_files"), 
    .progress = TRUE
)

## Ingest chunks into a KB database
kb <- BiocKB("biockb.duckdb", overwrite = TRUE)
purrr::map(chunks, insert_bioc_kb, kb = kb)
build_bioc_index(kb)
DBI::dbDisconnect(kb@store@con)
```

## Retrieve info from KB

```{r index}
# Connect to an existing store for querying
kb <- BiocKB("biockb.duckdb", read_only = TRUE)
retrieve_bioc_kb(kb, "How to perform differential expression analysis")$origin
retrieve_bioc_kb(kb, "How to submit a package to Bioconductor")$origin
```

# Session info

```{r sessionInfo, eval=TRUE}
sessionInfo()
```
