---
title: "Getting Started with BiocKB"
author: "Jacques Serizay"
date: "`r Sys.Date()`"
output:
    BiocStyle::html_document:
        toc: true
vignette: >
    %\VignetteIndexEntry{Getting Started with BiocKB}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Introduction

**BiocKB** builds and manages a RAG (Retrieval-Augmented Generation) knowledge
base from Bioconductor documentation. It provides a DuckDB-backed vector store
with chunked, embedded content from package vignettes, man pages, and other
Bioconductor resources.

# Installation

```{r install}
BiocManager::install("BiocKB")
```

# Store Lifecycle

The typical workflow follows: create → fetch → chunk → insert → index → connect.

## Create a Store

```{r create}
library(BiocKB)
store <- biockb_store_create(path = "biockb.duckdb")
```

## Fetch Package Resources

```{r fetch-pkgs}
# Download vignettes and man pages for selected packages
resources <- fetch_pkgs_resources(
    packages = c("DESeq2", "GenomicRanges"),
    destdir = "resources/"
)
```

## Chunk and Insert

```{r insert}
# Chunk files with AST-aware R parsing
chunks <- chunk_bioc_file("resources/DESeq2/vignettes/DESeq2.Rmd")

# Insert chunks into the store
biockb_store_insert(store, chunks)
```
## Index and Connect

```{r index}
# Build the embedding index
biockb_store_index(store)

# Connect to the store for querying
store <- biockb_store_connect(path = "biockb.duckdb")
```

# Additional Corpora

## Bioconductor Website

```{r website}
pages <- fetch_bioc_website()
```

## Support Site Posts

```{r support}
posts <- fetch_support_posts()
```

## biocViews Hierarchy

```{r views}
views <- fetch_bioc_views()
ingest_bioc_views(store, views)
```

# Session Info

```{r sessionInfo, eval=TRUE}
sessionInfo()
```
