---
title: "BiocAgentRAGDB: RAG Knowledge Base for BiocAgent"
author: "Bioconductor Core Team"
date: "`r Sys.Date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{BiocAgentRAGDB: RAG Knowledge Base for BiocAgent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Introduction

**BiocAgentRAGDB** provides a pre-built retrieval-augmented generation (RAG)
database containing the source code, vignettes, and metadata from all
Bioconductor packages (~3,000 packages). This database enables the `BiocAgent`
chat assistant to ground its LLM responses in real Bioconductor documentation.

The database is:

- **Hosted on ExperimentHub** for easy, versioned distribution
- **Cached locally via BiocFileCache** for fast repeated access
- **DuckDB-based** (via `ragnar`) with dual BM25 + vector similarity indexing

# Installation

```{r install}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("BiocAgentRAGDB")
```

# Fetching the Database

On first use, download the database from ExperimentHub:

```{r fetch}
library(BiocAgentRAGDB)

# Download and cache locally
db_path <- ragdb_fetch()

# Check where it's stored
ragdb_info()
```

The database is cached in `tools::R_user_dir("BiocAgentRAGDB", "data")` by
default. Subsequent calls to `ragdb_fetch()` return the cached copy instantly.

# Updating the Database

When a new version is released on ExperimentHub:

```{r update}
ragdb_update()
```

# Using with BiocAgent

The primary use case is as a knowledge source for `BiocAgent`:

```{r biocagent}
library(BiocAgent)

# Option 1: BiocAgent auto-detects the database
chat <- bioc_chat(ragdb = TRUE)

# Option 2: Explicit path
chat <- bioc_chat(ragdb = ragdb_path())

# The LLM now has access to Bioconductor documentation
chat$chat("How do I use DESeq2 for differential expression analysis?")
```

# Configuration

## Custom Cache Location

Override the default cache path:

```{r config}
# Via R option
options(BiocAgentRAGDB.path = "/data/shared/bioc_ragdb.duckdb")

# Via environment variable (e.g. in .Renviron)
# BIOCAGENT_RAGDB_PATH=/data/shared/bioc_ragdb.duckdb
```

# Database Contents

The RAG database is built from all Bioconductor packages and includes:

| Source | Content |
|--------|---------|
| `R/*.R` | Function source code and roxygen documentation |
| `vignettes/` | Vignettes (Rmd, qmd, Rnw, md) |
| `DESCRIPTION` | Package metadata (title, description, dependencies) |
| `NAMESPACE` | Exported/imported functions |

See `inst/scripts/build_ragdb.Rmd` in the package source for the complete
build pipeline.

# Session Info

```{r sessionInfo, eval=TRUE}
sessionInfo()
```
