---
title: "Building the BiocRAGDB from Bioconductor Package Sources"
author: "Bioconductor Core Team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Overview

This notebook documents the complete pipeline for building the BiocRAGDB
— a DuckDB-backed RAG (retrieval-augmented generation) database containing
the source code, vignettes, and metadata from all Bioconductor packages.

The resulting database file is uploaded to ExperimentHub for distribution via
the `BiocRAGDB` package.

```{r install-deps}
BiocManager::install(c("ragnar", "BiocPkgTools"))
```

# Clone package sources

All Bioconductor packages are hosted at `https://git.bioconductor.org/packages/`.
We use shallow clones (`--depth 1`) to minimize bandwidth and disk usage.

```{r setup-clone-dir}
library(BiocPkgTools)
library(ragnar)
clone_base <- "/pasteur/appa/scratch/jaseriza/BiocRAGDB"
files_dirs <- file.path(clone_base, "all_files")
db_path <- file.path(clone_base, "bioc_ragdb.duckdb")
all_files <- list.files(files_dirs, full.names = TRUE, recursive = TRUE)

clone_package <- function(pkg, base_dir, branch = "devel") {
    repo_url <- sprintf("https://git.bioconductor.org/packages/%s", pkg)
    dest <- file.path(base_dir, pkg)
    if (dir.exists(dest)) return(dest)
    result <- tryCatch(
        {
            system2(
                "git", c(
                    "clone", "--depth", "1",
                    "--branch", branch,
                    "--single-branch",
                    repo_url, dest
                ),
                stdout = TRUE, stderr = TRUE
            )
            dest
        },
        error = function(e) {
            warning(sprintf("Failed to clone %s: %s", pkg, e$message))
            NA_character_
        }
    )
    result
}

dir.create(clone_base, showWarnings = FALSE, recursive = TRUE)
pkg_tbl <- biocPkgList()
pkg_names <- pkg_tbl$Package
future::plan(future::multisession, workers = 8L)
future.apply::future_vapply(pkg_names, clone_package, character(1), base_dir = clone_base, future.seed = NULL, .progress = TRUE)
```

# List all files then extract and chunk them

```{r extract-helpers}
collect_package_files <- function(pkg_dir, out_dir) {
    pkg_name <- basename(pkg_dir)
    files <- character(0)

    # R/ source files
    r_dir <- file.path(pkg_dir, "R")
    if (dir.exists(r_dir)) {
        r_files <- list.files(
            r_dir, pattern = "\\.[Rr]$",
            full.names = TRUE, recursive = FALSE
        )
        files <- c(files, r_files)
    }

    # Vignettes
    vig_dir <- file.path(pkg_dir, "vignettes")
    if (dir.exists(vig_dir)) {
        vig_pattern <- paste0(
            "\\.(",
            paste(c("Rmd", "rmd", "qmd", "Rnw", "rnw", "md"), collapse = "|"),
            ")$"
        )
        vig_files <- list.files(
            vig_dir, pattern = vig_pattern,
            full.names = TRUE, recursive = TRUE
        )
        files <- c(files, vig_files)
    }

    # DESCRIPTION
    desc_file <- file.path(pkg_dir, "DESCRIPTION")
    if (file.exists(desc_file)) files <- c(files, desc_file)

    # NAMESPACE
    ns_file <- file.path(pkg_dir, "NAMESPACE")
    if (file.exists(ns_file)) files <- c(files, ns_file)

    # Now copy these files to the output directory, preserving relative paths
    if (length(files) == 0L) {
        warning(sprintf("No relevant files found for %s", pkg_name))
    } else {
        purrr::map(files, function(f) {
            rel_path <- sub(
                paste0(".*?/", pkg_name, "/"),
                paste0(pkg_name, "/"),
                f
            )
            dest_path <- file.path(files_dirs, rel_path)
            dir.create(dirname(dest_path), showWarnings = FALSE, recursive = TRUE)
            file.copy(f, dest_path, overwrite = TRUE)
        })
    }

}

pkg_dirs <- list.dirs(clone_base, full.names = TRUE, recursive = FALSE)
pkg_dirs <- pkg_dirs[pkg_dirs != file.path(clone_base, "all_files")]
purrr::map(pkg_dirs, collect_package_files, out_dir = files_dirs, .progress = TRUE)
all_files <- list.files(files_dirs, full.names = TRUE, recursive = TRUE)
```

# Process each file into chunks and insert into the ragnar store

```{r build-store}
## Define the embedding function
## Using Ollama with nomic-embed-text for a fully local build.
## THIS REQUIRES OLAMA TO BE INSTALLED AND THE MODEL TO BE AVAILABLE LOCALLY.
##        sh> ollama pull nomic-embed-text
store <- ragnar::ragnar_store_create(
    db_path,
    embed = function(x) {ragnar::embed_ollama(x, model = "nomic-embed-text")}, 
    embedding_size = 768L,
    overwrite = TRUE, 
    name = "BiocRAGDB", 
    title = "BiocRAGDB"
)
for (f in all_files[1:100]) {
    print(paste("Processing:", f))
    rel_path <- sub(paste0(files_dirs, "/"), "", f, fixed = TRUE)
    chunks <- BiocRAGDB::chunk_bioc_file(f, origin = rel_path)
    ragnar::ragnar_store_insert(store, chunks)
}
```

# Step 5: Build Indexes

After all chunks are inserted, build the full-text and vector similarity
indexes for efficient retrieval.

```{r build-indexes}
ragnar::ragnar_store_build_index(store)
message("Indexes built. Database ready at: ", db_path)
```

# Step 6: Validate the Database

Run some sanity checks on the built database.

```{r validate}
## Reconnect to verify
store <- ragnar::ragnar_store_connect(location = "/pasteur/appa/scratch/jaseriza/BiocRAGDB/bioc_ragdb.duckdb")

## Test a retrieval
results <- ragnar::ragnar_retrieve(
    store,
    "how can I esimate dropouts in single cell data?",
    top_k = 20
)
print(results[, c("origin", "doc_id", "text")])

## Test another query
results2 <- ragnar::ragnar_retrieve(
    store,
    "how to create a GRanges object",
    top_k = 5
)
print(results2[, c("origin", "doc_id", "text")])
```

# Step 7: Upload to ExperimentHub

The final database file is submitted to ExperimentHub following the standard
Bioconductor ExperimentHub resource submission process.

```{r eh-metadata}
## Generate the ExperimentHub metadata CSV
## This file goes in inst/extdata/metadata.csv of BiocRAGDB
metadata <- data.frame(
    Title = "BiocRAGDB: RAG Knowledge Base for BiocAgent",
    Description = paste(
        "DuckDB-backed RAG database containing chunked and embedded",
        "source code, vignettes, DESCRIPTION, and NAMESPACE files from",
        "all Bioconductor software packages. Built with ragnar."
    ),
    BiocVersion = as.character(BiocManager::version()),
    Genome = NA_character_,
    SourceType = "DuckDB",
    SourceUrl = "https://git.bioconductor.org/packages/",
    SourceVersion = format(Sys.Date(), "%Y%m%d"),
    Species = NA_character_,
    TaxonomyId = NA_integer_,
    Coordinate_1_based = NA,
    DataProvider = "Bioconductor",
    Maintainer = "Jacques Serizay <jacquesserizay@gmail.com>",
    RDataClass = "character",
    DispatchClass = "FilePath",
    RDataPath = "BiocRAGDB/bioc_ragdb.duckdb",
    stringsAsFactors = FALSE
)

## Write metadata for ExperimentHub submission
write.csv(
    metadata,
    file = "inst/extdata/metadata.csv",
    row.names = FALSE
)
```

```{r copy-for-upload}
## Copy the final database to the ExperimentHub staging area
final_path <- "bioc_ragdb.duckdb"
file.copy(db_path, final_path, overwrite = TRUE)
message(
    "Final database: ", final_path, "\n",
    "Size: ", round(file.info(final_path)$size / 1024^2, 1), " MB\n",
    "Submit this file to ExperimentHub."
)
```

# Session Info

```{r session-info, eval=TRUE}
sessionInfo()
```
