---
title: "Building the BiocRAGDB from Bioconductor Package Sources"
author: "Bioconductor Core Team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Overview

This notebook documents the complete pipeline for building the BiocRAGDB
— a DuckDB-backed RAG (retrieval-augmented generation) database containing
the source code, vignettes, and metadata from all Bioconductor packages.

The resulting database file is uploaded to ExperimentHub for distribution via
the `BiocRAGDB` package.

## Requirements

```{r install-deps}
BiocManager::install(c("ragnar", "BiocPkgTools"))
```

## Pipeline Summary

1. Enumerate all Bioconductor packages (~3,000)
2. Clone/shallow-fetch each package from `git.bioconductor.org`
3. Extract and chunk: `R/*.R`, `vignettes/*`, `DESCRIPTION`, `NAMESPACE`
4. Embed chunks and build the ragnar DuckDB store
5. Upload to ExperimentHub

# Step 1: Enumerate Bioconductor Packages

```{r list-packages}
library(BiocPkgTools)

## Fetch the full list of Bioconductor software packages
pkg_tbl <- biocPkgList()
pkg_names <- pkg_tbl$Package
message(sprintf("Found %d Bioconductor packages", length(pkg_names)))
```

# Step 2: Clone Package Sources

All Bioconductor packages are hosted at `https://git.bioconductor.org/packages/`.
We use shallow clones (`--depth 1`) to minimize bandwidth and disk usage.

```{r setup-clone-dir}
clone_base <- file.path(tempdir(), "bioc_sources")
dir.create(clone_base, showWarnings = FALSE, recursive = TRUE)
```

```{r clone-packages}
clone_package <- function(pkg, base_dir, branch = "devel") {
    repo_url <- sprintf(
        "https://git.bioconductor.org/packages/%s", pkg
    )
    dest <- file.path(base_dir, pkg)
    if (dir.exists(dest)) return(dest)
    result <- tryCatch(
        {
            system2(
                "git", c(
                    "clone", "--depth", "1",
                    "--branch", branch,
                    "--single-branch",
                    repo_url, dest
                ),
                stdout = TRUE, stderr = TRUE
            )
            dest
        },
        error = function(e) {
            warning(sprintf("Failed to clone %s: %s", pkg, e$message))
            NA_character_
        }
    )
    result
}

## Clone all packages (parallelized with future.apply if available)
future::plan(future::multisession, workers = 8L)
pkg_dirs <- future.apply::future_vapply(
    pkg_names, clone_package,
    character(1),
    base_dir = clone_base,
    future.seed = NULL
)

## Remove failed clones
pkg_dirs <- pkg_dirs[!is.na(pkg_dirs)]
message(sprintf("Successfully cloned %d packages", length(pkg_dirs)))
```

# Step 3: List all files then extract and chunk them

We extract four types of content per package:

| Source | File Pattern | Purpose |
|--------|-------------|---------|
| `R/*.R` | `R/` directory | Function code + roxygen docs |
| Vignettes | `vignettes/*.{Rmd,qmd,Rnw,md}` | Tutorials and workflows |
| `DESCRIPTION` | Root | Metadata: title, description, deps |
| `NAMESPACE` | Root | Exported/imported symbols |

```{r extract-helpers}
collect_package_files <- function(pkg_dir) {
    pkg_name <- basename(pkg_dir)
    files <- character(0)

    # R/ source files
    r_dir <- file.path(pkg_dir, "R")
    if (dir.exists(r_dir)) {
        r_files <- list.files(
            r_dir, pattern = "\\.[Rr]$",
            full.names = TRUE, recursive = FALSE
        )
        files <- c(files, r_files)
    }

    # Vignettes
    vig_dir <- file.path(pkg_dir, "vignettes")
    if (dir.exists(vig_dir)) {
        vig_pattern <- paste0(
            "\\.(",
            paste(c("Rmd", "rmd", "qmd", "Rnw", "rnw", "md"), collapse = "|"),
            ")$"
        )
        vig_files <- list.files(
            vig_dir, pattern = vig_pattern,
            full.names = TRUE, recursive = TRUE
        )
        files <- c(files, vig_files)
    }

    # DESCRIPTION
    desc_file <- file.path(pkg_dir, "DESCRIPTION")
    if (file.exists(desc_file)) files <- c(files, desc_file)

    # NAMESPACE
    ns_file <- file.path(pkg_dir, "NAMESPACE")
    if (file.exists(ns_file)) files <- c(files, ns_file)

    files
}

file_to_document <- function(file_path, pkg_name) {
    rel_path <- sub(
        paste0(".*?/", pkg_name, "/"),
        paste0(pkg_name, "/"),
        file_path
    )
    ext <- tolower(tools::file_ext(file_path))

    content <- tryCatch(
        paste(readLines(file_path, warn = FALSE), collapse = "\n"),
        error = function(e) return(NA_character_)
    )
    if (is.na(content) || nchar(content) == 0L) return(NULL)

    # Format as markdown with package/file context as headings
    lang <- switch(ext,
        r = "r",
        rmd = "rmd", qmd = "qmd", rnw = "rnw",
        md = "markdown",
        ""
    )
    md_text <- paste0(
        "# Package: ", pkg_name, "\n",
        "## File: ", rel_path, "\n\n",
        "```", lang, "\n",
        content, "\n",
        "```\n"
    )

    # Create a MarkdownDocument with origin metadata
    # The origin is preserved through chunking and into the store
    ragnar::MarkdownDocument(md_text, origin = file_path)
}

process_package_files <- function(pkg_dir) {
    pkg_name <- basename(pkg_dir)
    files <- collect_package_files(pkg_dir)

    if (length(files) == 0L) return(list())

    docs <- lapply(files, function(f) {
        tryCatch(
            file_to_document(f, pkg_name),
            error = function(e) {
                warning(
                    sprintf("Failed to process %s: %s", f, e$message)
                )
                NULL
            }
        )
    })
    docs <- Filter(Negate(is.null), docs)
    return(docs)
}

process_package_chunks <- function(files) {
    chunks_list <- lapply(files, function(f) {
        rel_path <- sub(
            paste0(".*?/", pkg_name, "/"),
            paste0(pkg_name, "/"),
            f
        )

        tryCatch(
            ragnar::markdown_chunk(f),
            error = function(e) {
                warning(
                    sprintf("Chunking failed for %s: %s", rel_path, e$message)
                )
                NULL
            }
        )
    })

    Filter(Negate(is.null), chunks_list)
}
```

# Step 4: Create the RAG Store and Ingest

```{r create-store}
## Define the embedding function
## Using Ollama with nomic-embed-text for a fully local build.
## THIS REQUIRES OLAMA TO BE INSTALLED AND THE MODEL TO BE AVAILABLE LOCALLY.
##        sh> ollama pull nomic-embed-text
library(ragnar)
db_path <- file.path(tempdir(), "bioc_ragdb.duckdb")
store <- ragnar::ragnar_store_create(db_path, embed = function(x) {ragnar::embed_ollama(x, model = "nomic-embed-text")})

for (i in seq_along(pkg_dirs)) {
    message(sprintf("[%d/%d] Processing %s...", i, n_pkgs, pkg_name))
    pkg_dir <- pkg_dirs[i]
    pkg_name <- basename(pkg_dir)
    pkg_files <- process_package_files(pkg_dir)
    chunks_list <- tryCatch(
        process_package_chunks(pkg_files),
        error = function(e) {
            warning(sprintf("Failed to process %s: %s", pkg_name, e$message))
            list()
        }
    )

    for (chunks in chunks_list) {
        ragnar::ragnar_store_insert(store, chunks)
    }
}
```

# Step 5: Build Indexes

After all chunks are inserted, build the BM25 full-text and vector similarity
indexes for efficient retrieval.

```{r build-indexes}
ragnar::ragnar_store_build_index(store)
message("Indexes built. Database ready at: ", db_path)
```

# Step 6: Validate the Database

Run some sanity checks on the built database.

```{r validate}
## Reconnect to verify
store <- ragnar::ragnar_store_connect(location = db_path)

## Test a retrieval
results <- ragnar::ragnar_retrieve(
    store,
    "differential expression analysis with DESeq2",
    top_k = 5
)
print(results[, c("origin", "doc_id", "text")])

## Test another query
results2 <- ragnar::ragnar_retrieve(
    store,
    "how to create a GRanges object",
    top_k = 5
)
print(results2[, c("origin", "doc_id", "text")])
```

# Step 7: Upload to ExperimentHub

The final database file is submitted to ExperimentHub following the standard
Bioconductor ExperimentHub resource submission process.

```{r eh-metadata}
## Generate the ExperimentHub metadata CSV
## This file goes in inst/extdata/metadata.csv of BiocRAGDB
metadata <- data.frame(
    Title = "BiocRAGDB: RAG Knowledge Base for BiocAgent",
    Description = paste(
        "DuckDB-backed RAG database containing chunked and embedded",
        "source code, vignettes, DESCRIPTION, and NAMESPACE files from",
        "all Bioconductor software packages. Built with ragnar."
    ),
    BiocVersion = as.character(BiocManager::version()),
    Genome = NA_character_,
    SourceType = "DuckDB",
    SourceUrl = "https://git.bioconductor.org/packages/",
    SourceVersion = format(Sys.Date(), "%Y%m%d"),
    Species = NA_character_,
    TaxonomyId = NA_integer_,
    Coordinate_1_based = NA,
    DataProvider = "Bioconductor",
    Maintainer = "Jacques Serizay <jacquesserizay@gmail.com>",
    RDataClass = "character",
    DispatchClass = "FilePath",
    RDataPath = "BiocRAGDB/bioc_ragdb.duckdb",
    stringsAsFactors = FALSE
)

## Write metadata for ExperimentHub submission
write.csv(
    metadata,
    file = "inst/extdata/metadata.csv",
    row.names = FALSE
)
```

```{r copy-for-upload}
## Copy the final database to the ExperimentHub staging area
final_path <- "bioc_ragdb.duckdb"
file.copy(db_path, final_path, overwrite = TRUE)
message(
    "Final database: ", final_path, "\n",
    "Size: ", round(file.info(final_path)$size / 1024^2, 1), " MB\n",
    "Submit this file to ExperimentHub."
)
```

# Session Info

```{r session-info, eval=TRUE}
sessionInfo()
```
